<!-- R Commander Markdown Template -->

SMDE - First Assignment
=======================

### Juan Pablo Royo Sales

### `r as.character(Sys.Date())`

```{r echo=FALSE}
# include this code chunk as-is to set options
knitr::opts_chunk$set(comment=NA, prompt=TRUE, out.width=750, fig.height=8, fig.width=8)
library(Rcmdr)
library(car)
library(RcmdrMisc)
```


```{r echo=FALSE}
# include this code chunk as-is to enable 3D graphs
library(rgl)
knitr::knit_hooks$set(webgl = hook_webgl)
```

## Loading data generated by Java FibFlagged method.

1. Load Data

```{r}
fibflagged <- read.table("/Users/juan/Projects/upc/master/smde/rng-j/sample.csv", 
    header = FALSE, sep = ",", na.strings = "NA", dec = ".", strip.white = TRUE)
```

2. Apply bins

``{r}
fibflagged$bins <- with(fibflagged, binVariable(V1, bins = 10, method = "intervals", 
    labels = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10")))
```

3. Generate Frequency

```{r}
fibflagged_trans <- as.data.frame(with(fibflagged, table(bins)))
```



## Generating uniform distribution with R

1. Load data

```{r}
distribution_r <- as.data.frame(matrix(runif(500 * 1, min = 0, max = 1), ncol = 1))
rownames(distribution_r) <- paste("sample", 1:500, sep = "")
colnames(distribution_r) <- "obs"
distribution_r <- within(distribution_r, {
    mean <- rowMeans(distribution_r[, 1:1])
})
```

2. Apply bins

```{r}
distribution_r$bins <- with(distribution_r, binVariable(obs, bins = 10, method = "intervals", 
    labels = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10")))
```

3. Generate Frequency

```{r}
distribution_r_trans <- as.data.frame(with(distribution_r, table(frequency)))
```

## Merge both frequencies R and FibFlagged

```{r}
merge_fibflagged_r_comb <- merge(fibflagged_trans, distribution_r_trans, all = TRUE, 
    by = "row.names")
rownames(merge_fibflagged_r_comb) <- merge_fibflagged_r_comb$Row.names
merge_fibflagged_r_comb$Row.names <- NULL
```

## Generate table for Chi test

```{r}
to_be_test_chi_r_fibflagged <- within(merge_fibflagged_r_comb, {
	bins <- NULL
	frequency <- NULL
})
```

## Run Chi Test

```{r}
test <- chisq.test(to_be_test_chi_r_fibflagged, correct = FALSE)
```

After this **test** has a p-value of **0.818475** con lo cual podemos deducir que la muestra es aceptable.



